Realizing temporal graphs from fastest travel times

### 2023-02-07 ICALP23-RealizingTemporalGraphs.tex
Changes to last section - ILP solution

@@

Let $G=(V,E)$ and $\Delta\in \mathbb{N}$, and let $\lambda: E \rightarrow \{1,2,\ldots,\Delta\}$ be an edge-labeling function that assigns to every edge of $G$ one of the labels from $\{1,\ldots,\Delta\}$. 
Then we denote by $(G,\lambda,\Delta)$ the \emph{$\Delta$-periodic temporal graph} $(G,\lambda_{\Delta})$, where for every edge $e\in E$ we have $\lambda_{\Delta}(e)=\{i\Delta + x : i\geq 0, x\in \lambda(e)\}$. 
In this case we call $\lambda$ a \emph{$\Delta$-periodic labeling} of $G$.
When it is clear from the context, we drop $\Delta$ form the notation and say that $(G,\lambda)$ is a $\Delta$-periodic temporal graph.

\begin{definition}[temporal path] \label{def:temporalPath+Duration}
Let $(G,\lambda)$ be a temporal graph. A \emph{temporal path} $(P,\lambda)$ from $u=u_0$ to $v=u_k$ in $(G,\lambda)$ is a sequence $(u_0 u_1,t_1),(u_1 u_2,t_2),\ldots,(u_{k-1} u_k,t_k)$, 
where $P=(u_0,u_1,\ldots,u_k)$ is a path in the underlying static graph $G$ (we usually say that $P$ is the underlying path of $(P,\lambda)$), 
$t_i\in \lambda(u_{i-1}u_i)$ for every $i=1,\ldots,k$, and $t_1<t_2<\ldots<t_k$. 
The \emph{duration} of this temporal path $(P,\lambda)$ from $u$ to $v$ is $d(P,\lambda)=t_k - t_1 + 1$.
A \emph{fastest} temporal path $(P,\lambda)$ from $u$ to $v$ in $(G,\lambda)$ is a temporal path from $u$ to $v$ with the smallest possible duration.
The duration of the \emph{fastest} temporal path from $u$ to $v$ is denoted by $d(u,v)$.
\end{definition}

Let $P$ be a path from $u$ to $v$ in $G$. 
We denote with $d(P,u,v)$ the duration of the fastest temporal path from $u$ to $v$ along path $P$. Note that $d(P,u,v) \geq d(u,v)$.

@@

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$ that we are looking for in the above problems. 
For simplicity of the presentation (and with a slight abuse of notation), we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as 
$D_{a,b}$, where $a=v_i$ and $b=v_j$. That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ instead of their numbering.
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda)$,
where we actually mean that $(P,\lambda)$ is a temporal path with the underlyg path $P$.

\begin{comment}
\begin{definition}[$D$-increasing temporal path]
Let $(G,\lambda,\Delta)$ be a $\Delta$-periodic temporal graph on $n$ vertices, and let $D$ be an $n\times n $ matrix of positive integers. 
Let $P$ be a temporal path from $u_0$ to $u_k$ on the vertices $u_0,u_1,\ldots,u_k$, and let $P^\textsc{ST}=(u_0,u_1,\ldots,u_k)$ be the underlying \emph{static} path in $G$.
Then $P$ is \emph{$D$-increasing} in $G$ if $0=D_{u_0,u_0}<D_{u_0,u_1}<D_{u_0,u_2}< \ldots < D_{u_0,u_k}$.
\end{definition}
\end{comment}

Let $(G,\lambda)$ be a temporal graph with  $n = |V(G)|$ and  $m=|E(G)|$.
Let $v \in V(G)$ be an arbitrary vertex of $G$, denote with $deg_{max}$ the maximum degree of all vertices in $V(G)$ and let $S_v$ be the set of time edges in $(G,\lambda)$, incident to $v$, \ie $S_v = \{(uv, \lambda(uv)) | uv \in E(G)\}$.
In their work Wu et~al.~\cite{Wu2016Efficient} provide an algorithm that computes the duration of the fastest temporal path from a source vertex $v$ to all other vertices in $G$ in
$O(n + m \log c)$ time, where $c = \min \{|S_v|, deg_{max}\}$.
From this result, we can deduce the following. 
\begin{lemma} \label{lemma:calculatingD}
Let $(G, \lambda, \Delta)$ be a $\Delta$-periodic temporal graph, where $\lambda$ assigns exactly $k$ labels to every edge of $G$. 
Denote with $deg_{max}$ the maximum degree of vertices in $G$, let $deg(v_i)$ be the degree of vertex $v_i \in V(G)$,
and let $c_{v_i} = \min \{ k \cdot deg(v_i), deg_{max}\}$.\\
To calculate the durations of all fastest temporal paths from a fixed vertex $v \in V(G)$ to all other vertices, it takes $O(n + m \log c_v)$ time. \\
To calculate the durations of the fastest temporal paths among all pairs of vertices we then need $O(n^2 + m \sum_{v \in V(G)} \log c_v)$ time.
\end{lemma}

Observe that $O(n + m \log c_v) = O(n + m \log deg_{max}) = O(n + m \log n) = O(n^2 \log n)$ and similarly
$O(n^2 + m \sum_{v \in V(G)} \log c_v) = O(n^2 + mn \log deg_{max}) = O(n + mn \log n) = O(n^3 \log n)$.
Since an instance $I$ of \kDeltaExact\ problem is an $n \times n$ matrix $D$, the running time of calculating one row of $D$ (\ie the fastest temporal paths from a single vertex to all others) takes $O(|I|\log \sqrt{|I|})$-time and calculating the whole matrix $D$ (\ie the fastest temporal paths among all pairs of vertices) takes $O(|I|^{3/2} \log \sqrt{|I|})$ time.
Therefore, from now on we say that calculating (a row of) $D$ takes polynomial time.

%\section{Results}

@@

    \begin{definition}[Travel delays]
    Let $(G, \lambda, \Delta)$ be a temporal graph satisfying conditions of problem \deltaExact.
    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
    More specifically:
    \begin{equation}\label{eq:def-VertexWaitingTime}
       \tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
    \end{equation}
    Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
    \end{definition}
    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

@@

\subparagraph{Matrix $D$.} We proceed with describing the matrix $D$, of durations of fastest paths. For a more convenient presentation, we use the notation $d(v,v'):= D_{v,v'}$. For all vertices $v,v'$ that are neighbors in $G$ we have that $d(v,v')=1$ and $d(v',v)=1$.

@@

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
    Note $S_{u,v} \neq S_{v,u}$.
\end{definition}
%From the definition it follows, that there are $O(k^2)$ segments in $G'$. \todo{True but does not follow directly just from the definition}
Observe that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments.
In the analysis of the algorithm we rely heavily on the fact that there are $O(k^2)$ segments in $G'$, 
which we prove later on (see~\cref{obs:FPT-k2segments}).
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

@@

\begin{lemma}\label{lemma:FPT-uv-Labelalledges}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1,v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a $u,v$ path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Moreover suppose also that $P$ is a segment.
    %Suppose also that no other vertex from $P$ is a vertex of interest, \ie $P \setminus \{u,v\} \cap U = \emptyset$.
    We can determine the labeling $\lambda$ of every edge in $P$ with respect to the label $\lambda(uv_2)$ of the first edge.
    %such that the labeling $\lambda$ respects $D'$.
\end{lemma}

@@

\todo[inline]{HM: the last sentence in \cref{lemma:FPT-uv-LabelAlmostalledges} is strange. it doesn't say anything. either give the specific restrictions or remove it.}

\begin{lemma}\label{lemma:FPT-uv-LabelAlmostalledges}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1, v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a $u,v$ path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Let $\ell_{u,v} \geq 1$ be the number of vertices of interest in $P$ different to $u,v$, namely $\ell_{u,v} = |\{P \setminus \{u,v \} \} \cap U |$.
    We can determine the labeling $\lambda$ of all but $\ell_{u,v}$ edges of $P$, with respect to the label $\lambda(uv_2)$ of the first edge,
    such that the labeling $\lambda$ respects the values from $D$.
    Moreover, for the labels of edges that cannot be determined precisely, we introduce some extra restrictions.
\end{lemma}

@@

\todo[inline]{HM: check whether we introduced the notation $d(u,v,P)$.}

@@

\subsection{Adding constraints and variables to the ILP}
We start by analyzing the case where we want to determine the labels on fastest temporal paths among vertices of interest.
We proceed in the following way.
Let $u,v \in U$ be two vertices of interest and let $P_{u,v}$ be the fastest temporal path from $u$ to $v$.
If $P_{u,v}$ is a segment we determine all the labels of edges of $P$, with respect to the label of the first edge (see \cref{lemma:FPT-uv-Labelalledges}).
%Note that since we determined labels of all edges, with respect to the first edge, we also know the exact values of waiting times at each inner vertex of $P_{u,v}$, when traversing the path.
In the case when $P_{u,v}$ is a sequence of $\ell$ segments, we determine all but $\ell - 1$ labels of edges of $P_{u,v}$, with respect to the label of the first edge (see \cref{lemma:FPT-uv-LabelAlmostalledges}).
%In both cases we introduce the equality constraint for $P_{u,v}$ (as in \cref{eq:FPT-equalityConstraint}).
%Since there are $\ell - 1$ edge labels that are not determined with respect to the label of the first edge, may be some  TODO from above??

After finishing the previous step for all pairs of vertices in $U$,
the edges of fastest temporal paths from $u$ to $v$, where $u,v \in U$, are determined with respect to the label of the first edge of each path.
Let us call these edges \emph{determined edges}, and the edges that were not determined with the respect to some other edge label up to now the \emph{not yet determined} edges.
If the fastest temporal path among two vertices $u,v \in U$ is just an edge $e$, then we treat it as being determined, since it gets assigned a label $\lambda(e)$ with respect to itself.
Note that the not yet determined edges are exactly the ones that are not a part of any fastest temporal path.
Let us first observe some properties of such edges.

\begin{lemma}
    Not yet determined edges form disjoint segments in $G'$, where each segment starts and ends in some vertex of interest, 
    \ie for every not yet determined path $P$ there exist $u,v \in U$, such that $P = (u, v_1, v_2, \dots, v_p=v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_{p-1} \in V(G') \setminus U$.
\end{lemma}
\begin{proof}
    We prove this claim in two steps.
    First we prove that all inner vertices of $P$, \ie vertices $v_i \in  S \setminus \{u,v\}$ are of degree exactly $2$ and are not vertices of interest. 
    They are of degree at least $2$ since they are on a path $P$ and are incident to two not yet determined edges in  $P$.
    They are of degree at most $2$ as if they were of a higher degree they would be vertices of interest and in this case we could split $P$ into shorter segments (subpaths).

    Second, suppose that $P$ does not start in a vertex of interest, therefore $u \notin U$. As $u$ is the starting vertex of the not yet determined path $P$, there must exist a vertex of interest $u'$ such that a path between $u$ and $u'$ is labeled.
    Since $u$ is not a vertex of interest, the only reason to label edges between $u'$ and $u$ would be if they were a part of some fastest temporal path among two vertices of interest and $u$ would be an inner vertex on that path.
    Therefore $u$ would have to be incident to two labeled edges (that are part of some fastest temporal path) and one not yet determined edge, but in this case the degree of $u$ would be at least $3$ and therefore $u$ would be a vertex of interest.

    From all of the above, it follows that $P$ is a not yet determined segment in $G'$.
\end{proof}

\begin{lemma}
    After determining the labels of fastest temporal paths between all pairs of vertices of interest,
    there can be at most one not yet determined segment among each pair of them.
    %among any pair of vertices of interest $u,v \in U$ there can be at most one undetermined segment $S = (u, v_1, v_2, \dots, v_p, v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_p \in V(G') \setminus U$.
\end{lemma}
\begin{proof}
   Suppose for the contradiction that after determining the labels of all fastest temporal paths between vertices of interest,
   there are two not yet determined segments among $u,v \in U$.
   Denote them with $S_{u,v}$ and $S'_{u,v}$. 
   Then $C = S_{u,v} \cup S_{u,v}'$ forms a cycle in $G'$. Since $G'$ admits no loops, the cycle $C$ is of length at least $3$.
   Since $C$ is a cycle, at least one of its edges $e \in E(C)$ has to be in the feedback edge set $F$, and therefore, by the definition of vertices of interest $U$, 
   both endpoints of $e$ have to be in $U$.
   If $e = uv$ then there are only two vertices of interest in $C$, but the label of edge $e$ is determined.
   And if $e \neq uv$ then there are at least $3$ vertices of interest in $C$.
   So $S_{u,v}$ and $S_{u,v}'$ cannot be both fully not yet determined.
\end{proof}
From the above it follows that
after we finish determining the labels for all fastest temporal paths among vertices of interest,
there is at most one fully determined segment among two vertices of interest and at most one undetermined segment.
Therefore the following holds.
\begin{corollary}\label{obs:FPT-k2segments}
    There are at most $O(k^2)$ segments in $G'$.
\end{corollary}

@@

W.l.o.g. assume that we have guessed that the fastest path from $v_i$ to $z_j$

@@

if and only if, this is a not yet determined edge in relation to every other segment $S_{u,v}$ and its fixed vertex $v_i$.

@@

At this point $G$ is a graph, where each edge $e$ has a value for its label $\lambda(e)$
that depends on (\ie is a function of) some other label $\lambda(f)$ of edge $f$,
or it depends on no other label.
We now describe how we create variables and start building our ILP instances. 
For every edge $e$ in $G'$ that is incident to a vertex of interest we create a variable $x_e$ that can have values from $ \{1, 2, \dots, \Delta\}$.
Besides that we create one variable for each edge that is still not yet determined on a segment.
Since each vertex of interest is incident to at most $k$ edges, and each segment has at most one extra not yet determined edge, we create $O(k^2)$ variables.
At the end we create our final guess
\begin{enumerate}[G-1.]
% counter for guesses continues
\setcounter{enumi}{\value{guesscounter}}
%%
    \item \label{FPT:guessallPermutations}
We guess the permutation of all $O(k^2)$ variables,
together with the relation of each variable to the labels of edges incident to these not yet determined edges.
Namely, for an edge $e$ that is not yet determined, we set its value to $x_e$ and check labels of all of its neighbors,
which are determined by some other label,
and variables of the not yet determined neighbors,
and guess if $x_e$ is smaller, equal or bigger than the labels of the edges of its neighbors.
So, for any two variables $x_e$ and $x_f$, we know if $x_e < x_f$ or $x_e = x_f$, or $x_e > x_f$,
and for any neighboring edge $g$ of $e$ we know if $x_e < \lambda(g)$ or 
$x_e = \lambda(g)$, or $x_e > \lambda(g)$.
This results in $O(k^2 !) = O(k^{2k^2}) = O(k^{k^3})$ guesses
and consequently
each of the ILP instances we created up to now is further split into $O(k^{k^3})$ new ones.
\end{enumerate}

We proceed with adding constraints to each of our ILP instances.
First we add all constraints for the labels of edges that we have determined up to now.
We then continue to iterate through all pairs of vertices and start adding equality (resp.~inequality) constraints for the fastest (resp.~non-fastest) temporal paths among them.

@@

\subsubsection*{\boldmath Fastest paths among $u,v$ where $u \in U$ and $v \in V$}
First we consider the case when $u,v$ are two vertices of interest.
This was the case that we partially studied at the beginning of the algorithm, that helped us determine certain labels of the graph.
We now iterate again through all fastest temporal paths from $u$ to $v$ and introduce an equality constraint for them.
We then continue 
through all of the paths from $u$ to $v$, and for every one, that is not the underlying path of a fastest temporal path, we add the inequality constraint.
There are $O(k^k)$ possible paths from $u$ to $v$ in $G'$, therefore in this step we introduce $O(k^{k^3})$ constraints.

\subsubsection*{\boldmath Fastest paths among $u,x$ where $u \in U$ and $x \in V(G') \setminus U$}
Next we continue with the case of the fastest path from $u$ to $x$, where $u \in U$ and $x \in V(G') \setminus U$.
From the guesses G-\ref{FPT:guess-uToSegmentz2} and G-\ref{FPT:guess-splitFromUtoAnotherSegment} it follows that we know the fastest temporal paths from $u$ to all vertices in a segment $S_{w,v}$.
In this case we create the equality constraint for the fastest path and
iterate through all other paths, for which we introduce the inequality constraints.
There are $O(k^k)$ possible paths of $u \leadsto w$ (resp.~$u \leadsto v$),
and a unique way how to extend those paths from $w$ (resp.~$v$) to reach $x$ in $S_{w,v}$.
Therefore we add $O(k^k)$ inequality constraints.

@@

\subsubsection*{\boldmath Fastest paths among $x,y$ where $x \in V(G') \setminus U$ and $y \in V(G') \setminus U$}
Next is the case of determining the fastest path from $x$ to $y$, where $x,y \in V(G') \setminus U$.
We have two options.
\begin{enumerate}[(i)]
    \item Vertices $x,y$ are in the same segment $S_{u,v} = (u,v_1,v_2, \dots, v_p, v)$. 
    If the length of $S_{u,v}$ is less than $4$ then we know what is the fastest path among vertices.
    Suppose now that $S_{u,v}$ is of length at least $5$. Then there are at most two not yet determined edges in $S_{u,v}$.

@@

\section{Ideas}
\begin{itemize}
    \item Symmetric distance matrix $D$: polytime?
    \item Use $|D|_\infty$ (maximum value in $D$) + max degree $\Delta$ as parameter $\rightarrow$ should give FPT
    \item FPT with feedback edge number
    \item use as parameter $\max_{i,j} |D_{i,j}-D_{j,i}|$
    \item approximation (additive or multiplicative), maybe with graph as input
\end{itemize}

\problemdef{\kDeltaExactLong\ (\kDeltaExact)}
{An $n \times n$ matrix $D$ of positive integers.}
{Does there exist a graph $G=(V,E)$ with vertices $\{v_1,\ldots,v_{n}\}$ 
and a $\Delta$-periodic labeling $\lambda: E \rightarrow \{1,2,\ldots,\Delta\}^k$ such that, 
for every $i,j$, the duration of the fastest temporal path from $v_i$ to $v_j$ in the $\Delta$-periodic temporal graph $(G,\lambda,\Delta)$ is equal to $D_{i,j}$.}

>>>>

Due to space constraints, proofs of results marked with $\star$ are (partially) deferred to Appendix.

@@

Let $G=(V,E)$ and $\Delta\in \mathbb{N}$, and let $\lambda: E \rightarrow \{1,2,\ldots,\Delta\}$ be an edge-labeling function that assigns to every edge of $G$ one of the labels from $\{1,\ldots,\Delta\}$. 
Then we denote by $(G,\lambda,\Delta)$ the \emph{$\Delta$-periodic temporal graph} $(G,\lambda_{\Delta})$, where for every edge $e\in E$ we have $\lambda_{\Delta}(e)=\{i\Delta + x : i\geq 0, x\in \lambda(e)\}$. 
In this case we call $\lambda$ a \emph{$\Delta$-periodic labeling} of $G$.
When it is clear from the context, we drop $\Delta$ form the notation and say that $(G,\lambda)$ is a ($\Delta$-periodic) temporal graph.

\begin{definition}[temporal path] \label{def:temporalPath+Duration}
Let $(G,\lambda)$ be a temporal graph. A \emph{temporal path} $(P,\lambda)$ from $u=u_0$ to $v=u_k$ in $(G,\lambda)$ is a sequence $(u_0 u_1,t_1),(u_1 u_2,t_2),\ldots,(u_{k-1} u_k,t_k)$, 
where $P=(u_0,u_1,\ldots,u_k)$ is a path in the underlying static graph $G$ (we usually say that $P$ is the underlying path of $(P,\lambda)$), 
$t_i\in \lambda_\Delta(u_{i-1}u_i)$ for every $i=1,\ldots,k$, and $t_1<t_2<\ldots<t_k$. 
The \emph{duration} of this temporal path $(P,\lambda)$ from $u$ to $v$ is $d(P,\lambda)=t_k - t_1 + 1$.
A \emph{fastest} temporal path $(P,\lambda)$ from $u$ to $v$ in $(G,\lambda)$ is a temporal path from $u$ to $v$ with the smallest possible duration.
The duration of the \emph{fastest} temporal path from $u$ to $v$ is denoted by $d(u,v)$.
\end{definition}

Let $P$ be a path from $u$ to $v$ in $G$. 
We denote with $d(u,v,P)$ the duration of the fastest temporal path from $u$ to $v$ along path $P$. Note that $d(u,v,P) \geq d(u,v)$.

@@

Let $v_1,v_2,\ldots,v_n$ be the $n$ vertices of the graph $G$ that we are looking for in the above problems. 
For simplicity of the presentation (and with a slight abuse of notation), we refer during the paper to the entry $D_{i,j}$ of the matrix $D$ as 
$D_{a,b}$, where $a=v_i$ and $b=v_j$. That is, we put as indices of the matrix $D$ the corresponding vertices of $G$ instead of their numbering.
Many times we also refer to a path $P=(u=v_1, v_2, \dots, v_p=v)$ from $u$ to $v$ in $G$,
as a temporal path in $(G,\lambda)$,
where we actually mean that $(P,\lambda)$ is a temporal path with the underlying path $P$.

In their work Wu et~al.~\cite{Wu2016Efficient} provide an algorithm that computes the duration of the fastest temporal path from a source vertex $v$, in a temporal graph $(G,\lambda)$, to all other vertices,
that runs in polynomial-time.
Using their algorithm, given a temporal graph $(G,\lambda)$, one can determine the whole matrix $D$ in polynomial-time,
where $D$ consists of durations of fastest temporal paths among all pairs of vertices in $(G,\lambda)$.

\begin{comment}
Let $(G,\lambda)$ be a temporal graph with  $n = |V(G)|$ and  $m=|E(G)|$.
Let $v \in V(G)$ be an arbitrary vertex of $G$, denote with $deg_{max}$ the maximum degree of all vertices in $V(G)$ and let $S_v$ be the set of time edges in $(G,\lambda)$, incident to $v$, \ie $S_v = \{(uv, \lambda(uv)) | uv \in E(G)\}$.
In their work Wu et~al.~\cite{Wu2016Efficient} provide an algorithm that computes the duration of the fastest temporal path from a source vertex $v$ to all other vertices in $G$ in
$O(n + m \log c)$ time, where $c = \min \{|S_v|, deg_{max}\}$.
From this result, we can deduce the following. 
\begin{lemma} \label{lemma:calculatingD}
Let $(G, \lambda, \Delta)$ be a $\Delta$-periodic temporal graph, where $\lambda$ assigns exactly $k$ labels to every edge of $G$. 
Denote with $deg_{max}$ the maximum degree of vertices in $G$, let $deg(v_i)$ be the degree of vertex $v_i \in V(G)$,
and let $c_{v_i} = \min \{ k \cdot deg(v_i), deg_{max}\}$.\\
To calculate the durations of all fastest temporal paths from a fixed vertex $v \in V(G)$ to all other vertices, it takes $O(n + m \log c_v)$ time. \\
To calculate the durations of the fastest temporal paths among all pairs of vertices we then need $O(n^2 + m \sum_{v \in V(G)} \log c_v)$ time.
\end{lemma}

Observe that $O(n + m \log c_v) = O(n + m \log deg_{max}) = O(n + m \log n) = O(n^2 \log n)$ and similarly
$O(n^2 + m \sum_{v \in V(G)} \log c_v) = O(n^2 + mn \log deg_{max}) = O(n + mn \log n) = O(n^3 \log n)$.
Since an instance $I$ of \kDeltaExact\ problem is an $n \times n$ matrix $D$, the running time of calculating one row of $D$ (\ie the fastest temporal paths from a single vertex to all others) takes $O(|I|\log \sqrt{|I|})$-time and calculating the whole matrix $D$ (\ie the fastest temporal paths among all pairs of vertices) takes $O(|I|^{3/2} \log \sqrt{|I|})$ time.
Therefore, from now on we say that calculating (a row of) $D$ takes polynomial time.
\end{comment}
%\section{Results}

@@

    \begin{definition}[Travel delays]
    Let $(G, \lambda, \Delta)$ be a temporal graph satisfying conditions of \deltaExact.
    Let $e_1=uv$ and $e_2=vz$ be two incident edges in $G$ with $e_1 \cap e_2 = v$.
    We define the \emph{travel delay} from $u$ to $z$ at vertex $v$, denoted with $\tau_v^{uz}$,
    as the difference of the labels of $e_2$ and $e_1$, where we subtract the value of the label of $e_1$ from the label of $e_2$,  modulo~$\Delta$.
    More specifically:
    \begin{equation}\label{eq:def-VertexWaitingTime}
       \tau_v^{uz} \equiv \lambda (e_2) - \lambda(e_1) \pmod \Delta.
    \end{equation}
    Similarly, $\tau_v^{zu} \equiv \lambda (e_1) - \lambda(e_2) \pmod \Delta$.
    \end{definition}
    Intuitively, the value of $\tau_v^{uz}$ represents how long a temporal path waits at vertex $v$ when first taking edge $e_1=uv$ and then edge $e_2 = vz$.

@@

\subparagraph{\boldmath Duration Matrix $D$.} We proceed with describing the matrix $D$, of durations of fastest paths. For a more convenient presentation, we use the notation $d(v,v'):= D_{v,v'}$. For all vertices $v,v'$ that are neighbors in $G$ we have that $d(v,v')=1$ and $d(v',v)=1$.

@@

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
    Note $S_{u,v} \neq S_{v,u}$.
\end{definition}
%From the definition it follows, that there are $O(k^2)$ segments in $G'$. \todo{True but does not follow directly just from the definition}
Observe that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments. Furthermore, since we have at most $4k$ interesting vertices in $G'$, we can deduce the following.
\begin{corollary}\label{obs:FPT-k2segments}
    There are at most $O(k^2)$ segments in $G'$.
\end{corollary}
In the analysis of the algorithm we rely heavily on the fact that there are $O(k^2)$ segments in $G'$.
%, which we prove later on (see~\cref{obs:FPT-k2segments}).
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

@@

\begin{lemma}\label{lemma:FPT-uv-Labelalledges}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1,v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Moreover suppose also that $P$ is a segment.
    %Suppose also that no other vertex from $P$ is a vertex of interest, \ie $P \setminus \{u,v\} \cap U = \emptyset$.
    We can determine the labeling $\lambda$ of every edge in $P$ with respect to the label $\lambda(uv_2)$ of the first edge.
    %such that the labeling $\lambda$ respects $D'$.
\end{lemma}

@@

%\todo[inline]{HM: the last sentence in \cref{lemma:FPT-uv-LabelAlmostalledges} is strange. it doesn't say anything. either give the specific restrictions or remove it.}

\begin{lemma}\label{lemma:FPT-uv-LabelAlmostalledges}
    Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1, v_2, \dots, v_p = v)$, where $p \geq 2$, 
    is a path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
    Let $\ell_{u,v} \geq 1$ be the number of vertices of interest in $P$ different to $u,v$, namely $\ell_{u,v} = |\{P \setminus \{u,v \} \} \cap U |$.
    We can determine the labeling $\lambda$ of all but $\ell_{u,v}$ edges of $P$, with respect to the label $\lambda(uv_2)$ of the first edge,
    such that the labeling $\lambda$ respects the values from $D$.
%    Moreover, for the labels of edges that cannot be determined precisely, we introduce some extra restrictions.
\end{lemma}

@@

@@

\subsection{Adding constraints and variables to the ILP}
We start by analyzing the case where we want to determine the labels on fastest temporal paths among vertices of interest.
We proceed in the following way.
Let $u,v \in U$ be two vertices of interest and let $P_{u,v}$ be the fastest temporal path from $u$ to $v$.
If $P_{u,v}$ is a segment we determine all the labels of edges of $P_{u,v}$, with respect to the label of the first edge (see \cref{lemma:FPT-uv-Labelalledges}).
%Note that since we determined labels of all edges, with respect to the first edge, we also know the exact values of waiting times at each inner vertex of $P_{u,v}$, when traversing the path.
In the case when $P_{u,v}$ is a sequence of $\ell$ segments, we determine all but $\ell - 1$ labels of edges of $P_{u,v}$, with respect to the label of the first edge (see \cref{lemma:FPT-uv-LabelAlmostalledges}).
We call these $\ell - 1$ edges, \emph{partially determined} edges.
After repeating this step for all pairs of vertices in $U$,
the edges of fastest temporal paths from $u$ to $v$, where $u,v \in U$, are determined with respect to the label of the first edge of each path,
or are partially determined.
If the fastest temporal path among two vertices $u,v \in U$ is just an edge $e$, then we treat it as being determined, since it gets assigned a label $\lambda(e)$ with respect to itself.
All other edges in $G'$ %that were not determined with the respect to some other edge label up to now, or are not partially determined, 
are called the \emph{not yet determined} edges.
Note that the not yet determined edges are exactly the ones that are not a part of any fastest temporal path.

@@

W.l.o.g.\ assume that we have guessed that the fastest path from $v_i$ to $z_j$

@@

if and only if this is a not yet determined edge in relation to every other segment $S_{u,v}$ and its fixed vertex $v_i$.

@@

At this point $G$ is a graph, where each edge $e$ has a value for its label $\lambda(e)$
that depends on (\ie is a function of) some other label $\lambda(f)$ of edge $f$,
or it depends on no other label.
We now describe how we create variables and start building our ILP instances. 
For every edge $e$ in $G'$ that is incident to a vertex of interest we create a variable $x_e$ that can have values from $ \{1, 2, \dots, \Delta\}$.
Besides that we create one variable for each edge that is still not yet determined on a segment.
Since each vertex of interest is incident to at most $k$ edges, and each segment has at most one extra not yet determined edge, we create $O(k^2)$ variables.
At the end we create our final guess.
\begin{enumerate}[G-1.]
% counter for guesses continues
\setcounter{enumi}{\value{guesscounter}}
%%
    \item \label{FPT:guessallPermutations}
We guess the permutation of all $O(k^2)$ variables,
together with the relation of each variable to the labels of edges incident to these not yet determined edges.
Namely, for an edge $e$ that is not yet determined, we set its value to $x_e$ and check labels of all of its neighbors,
which are determined by some other label,
and variables of the not yet determined neighbors,
and guess if $x_e$ is smaller, equal or bigger than the labels of the edges of its neighbors.
So, for any two variables $x_e$ and $x_f$, we know if $x_e < x_f$ or $x_e = x_f$, or $x_e > x_f$,
and for any neighboring edge $g$ of $e$ we know if $x_e < \lambda(g)$ or 
$x_e = \lambda(g)$, or $x_e > \lambda(g)$.
This results in $O(k^2 !) = O(k^{2k^2}) = O(k^{k^3})$ guesses
and consequently
each of the ILP instances we created up to now is further split into $O(k^{k^3})$ new ones.
\end{enumerate}

We proceed with adding constraints to each of our ILP instances.
First we add all constraints for the labels of edges that we have determined up to now.
We then continue to iterate through all pairs of vertices and start adding equality (resp.~inequality) constraints for the fastest (resp.~not necessarily fastest) temporal paths among them.

@@

\subparagraph{\boldmath Fastest paths among $u,v$ where $u \in U$ and $v \in V$.}
First we consider the case when $u,v$ are two vertices of interest.
This was the case that we partially studied at the beginning of the algorithm, that helped us determine certain labels of the graph.
We now iterate again through all fastest temporal paths from $u$ to $v$ and introduce an equality constraint for them.
We then continue 
through all of the paths from $u$ to $v$, and for every one, that is not the underlying path of a fastest temporal path, we add the inequality constraint.
There are $O(k^k)$ possible paths from $u$ to $v$ in $G'$, therefore in this step we introduce $O(k^{k^3})$ constraints.

\subparagraph{\boldmath Fastest paths among $u,x$ where $u \in U$ and $x \in V(G') \setminus U$.}
Next we continue with the case of the fastest path from $u$ to $x$, where $u \in U$ and $x \in V(G') \setminus U$.
From the guesses G-\ref{FPT:guess-uToSegmentz2} and G-\ref{FPT:guess-splitFromUtoAnotherSegment} it follows that we know the fastest temporal paths from $u$ to all vertices in a segment $S_{w,v}$.
In this case we create the equality constraint for the fastest path and
iterate through all other paths, for which we introduce the inequality constraints.
There are $O(k^k)$ possible paths of $u \leadsto w$ (resp.~$u \leadsto v$),
and a unique way how to extend those paths from $w$ (resp.~$v$) to reach $x$ in $S_{w,v}$.
Therefore we add $O(k^k)$ inequality constraints.

@@

\subparagraph{\boldmath Fastest paths among $x,y$ where $x \in V(G') \setminus U$ and $y \in V(G') \setminus U$.}
Next is the case of determining the fastest path from $x$ to $y$, where $x,y \in V(G') \setminus U$.
We have two options.
\begin{enumerate}[(i)]
    \item Vertices $x,y$ are in the same segment $S_{u,v} = (u,v_1,v_2, \dots, v_p, v)$. 
    If the length of $S_{u,v}$ is less than $4$ then we know what is the fastest path among vertices.
    Suppose now that $S_{u,v}$ is of length at least $5$. Then there are at most two not yet determined edges in $S_{u,v}$.

@@
